Just Add IFC! A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/concrete.tex
rephase of some paragraphs

If the operational rules are adapted na\"ively to this new setting,
then non-interference can be violated.  In particular, it is not
sound to share references to a heap cell between different tasks.
Otherwise, information can be trivially leaked by copying
secret data into a reference in one task, and then
reading it again by a another task with a lower label.
What we would like is a way of characterizing safe modifications to
the semantics which preserve non-interference.
The intention of our single heap implementation was to be more efficient
by using one global heap, but conceptually maintain isolation between
task by not allowing sharing of references between them.
This intuition of having a different (potentially more efficient)
concrete semantics that behave like the abstract semantics
can be formalized by the following definition:

@@

\Red{the next paragraph needs some rephrasing. We should say that if we 
condense everything into one environment, then we need to ``simulate'' 
the different policies. A bit more intro}
A na\"ive attempt at our single heap semantics
will not allow us to construct an isomorphism
with the abstract language |specLangML alpha|, and instead highlights
where things will go wrong.
To successfully construct an isomorphism, our concrete semantics need
to mirror the behavior of the abstract semantics, most importantly
in the case where tasks interact.  Namely, at a fork and when sending
or receiving messages, we take an expression from one task and
execute it in the context of another task.  In the abstract semantics
the fork rule is parameterized by the strategy |klone|, which defines
what heap the new task should execute in.  \Red{In Section~\ref{sec:retrofit}
we have motivated two choices}, namely the identity (to effectively
clone the heap), or to start with an empty heap.
In the latter case, this effectively means that existing addresses
in the forked expression |ie| are no longer valid and the
task gets stuck (and then removed by \textsc{I-noStep}).
\todo{}{Is this actually true?  Currently the semantics are unclear.}
To get the isomorphic behavior in the concrete semantics with a
single heap, we would have to change the forked expression |ie| in
such a way that all addresses get rewritten to fresh addresses
that are guaranteed to not occur in the global heap.  Similarly,
sending a memory address should take a similar approach, as addresses
for sent references will not exist in the receiving task.

If |klone| is the identity function, then the abstract semantics
effectively clone the heap on a fork.  To get an isomorphic
behavior in the concrete, we would have to create a copy of everything
reachable from the forked expression (not just the values for
the memory addresses in the expression), and replace all addresses
correspondingly.  Even worse, the behavior of sending a memory address
now depends on whether that address existed at the time the receiving
task was forked;  if it did, then the address should be rewritten to the
new one, otherwise to a fresh address.
Such a complicated behavior is contradictory to our initial motivation
of implementing a single heap for efficiency.  But also the
behavior of the simpler case where |klone| returns an empty heap is
undesirable:  We rewrite addresses to fresh addresses, which will
result in a stuck task whenever such an address is dereferenced.

@@

Again, we are modifying the IFC language semantics in a non-trivial way,
which begs the question whether non-interference is preserved.
Restricting the transitions in an information flow control
system in general does not preserve non-interference as the following
minimal example illustrates. \Red{Ale: I am not sure that the example is clear,
  too abstract, at least, I don't get it. Propose to remove it, and just say 
the top level idea: ``the restriction should not be based on secret data!''} 
Consider a trivial state machine as shown
in Figure~\ref{fig:trivial-sm}, whose second projection is classified
secret.  With all three transitions, this language fulfills
termination sensitive non-interference.  Furthermore, if the
dashed transition
is removed, the language continues to satisfy TSNI.  However, if any
solid line is removed, the language fails TSNI.

\begin{figure}
  States: $(x,y)$ for $x,y \in \{0,1\}$ \\
  Erasure function: $f(x,y) = (x,\bullet)$

  \begin{center}\begin{tikzpicture}[node distance=2cm, auto]
    \node (A) {$(1,1)$};
    \node (B) [right of=A] {$(1,0)$};
    \node (C) [below of=A] {$(0,1)$};
    \node (D) [right of=C] {$(0,0)$};
    \draw[->] (A) to node {} (B);
    \draw[->] (C) to node {} (D);
    \draw[->, dashed] (B) to node {} (D);
    \end{tikzpicture}\end{center}

  \label{fig:trivial-sm}
  \caption{A trivial state machine}
\end{figure}

To remedy this issue, we give a definition that gives a condition
under which it is safe to restrict the abstract IFC language
|specLang alpha targetLang|, such that non-interference is preserved.

>>>>

If the operational rules are adapted na\"ively to this new setting,
then non-interference can be violated.  In particular, it is not
sound to share references to a heap cell between different tasks.
Otherwise, information can be trivially leaked by copying
secret data into a reference in one task, and then
reading it again by a another task with a lower label.
What we would like is a way of characterizing safe modifications to
the semantics which preserve non-interference.
The intention of our single heap implementation was to be more efficient
by using one global heap, but \emph{conceptually maintain isolation between
task by not allowing sharing of references between them}.
This intuition of having a different (potentially more efficient)
concrete semantics that behave like the abstract semantics
can be formalized by the following definition:

@@

To obtain an isomorphism with our single heap semantics, we need to mimic the
behavior of several heaps into one. The interesting cases are when forking and
sending or receiving messages. These rules (see Section~\ref{sec:retrofit}) are
parameterized by the strategy |klone|, which defines what heap the new task
should execute with.  We have motivated two choices, namely the identity (to
effectively clone the heap), or to start with an empty heap.

In the first choice, the concrete semantics should essentially implement several
heaps into one. For that, it creates a copy of everything reachable from the
forked expression (not just the values for the memory addresses in the
expression \Red{Ale:why not only the addresses?}), and replace all addresses
correspondingly.  behavior of sending a memory address now depends on whether
that address existed at the time the receiving task was forked; if it did, then
the address should be rewritten to the new one, otherwise to a fresh one.

In the latter case, this effectively means that the concrete semantics must
change the forked expression |ie| in such a way that all addresses get rewritten
to fresh addresses that are guaranteed to not occur in the global heap. As a
consequence, existing addresses in the forked expression |ie| are no longer
valid and the task gets stuck (and then removed by \textsc{I-noStep}).
%\todo{}{Is this actually true?  Currently the semantics are unclear.}
%To get the isomorphic behavior in the concrete semantics with a
%single heap, we would have to change the forked expression |ie| in
%such a way that all addresses get rewritten to fresh addresses
%that are guaranteed to not occur in the global heap.  
Similarly, sending a memory address should take a similar approach, as addresses
for sent references will not exist in the receiving task.

% A na\"ive attempt at our single heap semantics
% will not allow us to construct an isomorphism
% with the abstract language |specLangML alpha|, and instead highlights
% where things will go wrong.
% To successfully construct an isomorphism, our concrete semantics need
% to mirror the behavior of the abstract semantics, most importantly
% in the case where tasks interact.  Namely, at a fork and when sending
% or receiving messages, we take an expression from one task and
% execute it in the context of another task.  In the abstract semantics
% the fork rule is parameterized by the strategy |klone|, which defines
% what heap the new task should execute in.  \Red{In Section~\ref{sec:retrofit}
% we have motivated two choices}, namely the identity (to effectively
% clone the heap), or to start with an empty heap.
% In the latter case, this effectively means that existing addresses
% in the forked expression |ie| are no longer valid and the
% task gets stuck (and then removed by \textsc{I-noStep}).
% \todo{}{Is this actually true?  Currently the semantics are unclear.}
% To get the isomorphic behavior in the concrete semantics with a
% single heap, we would have to change the forked expression |ie| in
% such a way that all addresses get rewritten to fresh addresses
% that are guaranteed to not occur in the global heap.  Similarly,
% sending a memory address should take a similar approach, as addresses
% for sent references will not exist in the receiving task.

% If |klone| is the identity function, then the abstract semantics
% effectively clone the heap on a fork.  To get an isomorphic
% behavior in the concrete, we would have to create a copy of everything
% reachable from the forked expression (not just the values for
% the memory addresses in the expression), and replace all addresses
% correspondingly.  Even worse, the behavior of sending a memory address
% now depends on whether that address existed at the time the receiving
% task was forked;  if it did, then the address should be rewritten to the
% new one, otherwise to a fresh address.
% Such a complicated behavior is contradictory to our initial motivation
% of implementing a single heap for efficiency.  But also the
% behavior of the simpler case where |klone| returns an empty heap is
% undesirable:  We rewrite addresses to fresh addresses, which will
% result in a stuck task whenever such an address is dereferenced.

@@

Generally speaking, we are modifying the IFC language semantics in a non-trivial
way, which begs the question whether non-interference is preserved.  With this
in mind, we device a general condition under which restrictions on transitions
preserve non-interference. We require semantic rules to get restricted only
based on information observable by the task triggering them. In that manner, the
restriction does not depend on confidential information---a possible risk to
jeopardize security. Below, the formal definition of such conditions for the
abstract IFC language |specLang alpha targetLang|.

%, such that non-interference is preserved.

%system in general does not preserve non-interference as the following
%minimal example illustrates.

% Again, we are modifying the IFC language semantics in a non-trivial way,
% which begs the question whether non-interference is preserved.
% Restricting the transitions in an information flow control
% system in general does not preserve non-interference as the following
% minimal example illustrates. \Red{Ale: I am not sure that the example is clear,
%   too abstract, at least, I don't get it. Propose to remove it, and just say 
% the top level idea: ``the restriction should not be based on secret data!''} 
% Consider a trivial state machine as shown
% in Figure~\ref{fig:trivial-sm}, whose second projection is classified
% secret.  With all three transitions, this language fulfills
% termination sensitive non-interference.  Furthermore, if the
% dashed transition
% is removed, the language continues to satisfy TSNI.  However, if any
% solid line is removed, the language fails TSNI.

% \begin{figure}
%   States: $(x,y)$ for $x,y \in \{0,1\}$ \\
%   Erasure function: $f(x,y) = (x,\bullet)$

%   \begin{center}\begin{tikzpicture}[node distance=2cm, auto]
%     \node (A) {$(1,1)$};
%     \node (B) [right of=A] {$(1,0)$};
%     \node (C) [below of=A] {$(0,1)$};
%     \node (D) [right of=C] {$(0,0)$};
%     \draw[->] (A) to node {} (B);
%     \draw[->] (C) to node {} (D);
%     \draw[->, dashed] (B) to node {} (D);
%     \end{tikzpicture}\end{center}

%   \label{fig:trivial-sm}
%   \caption{A trivial state machine}
% \end{figure}

%To remedy this issue,
